---
title: "Week 5 Storytelling Graphs_sc"
author: "Elizabeth Delmelle"
date: "2025-09-18"
output: html_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(lemon)
library(ggtext)
library(tidycensus)
library(lubridate)
library(dplyr)
library(grid)
source("helper_functions.R")
source("theme/theme_swd.R")
theme_set(theme_swd())
```

The first graph we are going to create is a scatterplot

```{r figure 206 pre_processing}

df <- read_csv(file.path("data", "FIG0206-7.csv")) %>% 
  mutate(`Cost Per Mile` = as.numeric(str_remove_all(`Cost Per Mile`,"\\$"))) 

avg_miles <- mean(df$`Miles Driven`)
avg_cost <- mean(df$`Cost Per Mile`)
```

Now to the plot

```{r scatterplot 1}
pt <- ggplot(df) + 
  geom_point(aes(x = `Miles Driven`,y = `Cost Per Mile`), color = GRAY9, size = 2.5) +
  scale_y_continuous(breaks = seq(0,3.5,.5), limits = c(0,3), labels = scales::dollar_format()) + 
  scale_x_continuous(limits = c(0,4000), label = scales::comma) +
  geom_point(x = avg_miles, y = avg_cost, size=3.5) + 
  geom_text(x = avg_miles, y = avg_cost, label = "AVG", hjust = 1.25) +  
  labs(title = "Cost per mile by miles driven", x = "Miles driven per month", y = "Cost per mile") +
  coord_capped_cart(bottom = "right", left = "top")
  
  pt %>% 
  save_and_show_plot(width = 5, height = 4, "FIG0206.png")
```

Here we are going to snaz it up and make points above the average one color

```{r fig207 pre_process}
df <- read_csv(file.path("data", "FIG0206-7.csv")) %>% 
  mutate(`Cost Per Mile` = parse_number(`Cost Per Mile`),
         avg_miles = mean(`Cost Per Mile`, na.rm = TRUE),
         avg_cost = mean(`Cost Per Mile`, na.rm = TRUE),
         color = if_else(`Cost Per Mile` < avg_miles, GRAY9, ORANGE1))

avg_miles <- mean(df$`Miles Driven`)
avg_cost <- mean(df$`Cost Per Mile`)
```

Now we will make the graph

```{r figure207}
pt <- ggplot(df) + 
  geom_point(aes(x=`Miles Driven`,y=`Cost Per Mile`, color = color), size = 3) +
  scale_color_identity() + 
  scale_y_continuous(breaks = seq(0,3,.5), limits = c(0,3),labels = scales::dollar_format()) + 
  scale_x_continuous(limits = c(0,4000), label = scales::comma) + 
  geom_hline(yintercept = avg_cost, linetype = "longdash") + 
  geom_point(x = avg_miles,y=avg_cost,size=3.5) + 
  geom_label(x = avg_miles, y = avg_cost, label = "AVG", hjust = 1.25, label.size = 0) +  
  labs(title = "Cost per mile by miles driven", x = "Miles driven per month", y = "Cost per mile") +
  coord_capped_cart(bottom = "right", left = "top")

pt %>% 
  save_and_show_plot(width = 4.5, height = 4.2, "FIG0207.png")
```
Challenge time!
get some census data and make a modified scatterplot. Add two lines and highlight counties that are above the median income and above the median age.

```{r getcensus}
library(tidycensus)

pa_wide_b <- get_acs(
  geography = "county",
  state = "Pennsylvania",
  variables = c(medinc = "B19013_001",
                medage = "B01002_001"),
  output = "wide",
  year = 2020
)

pa_wide <- pa_wide_b %>%
  mutate(avg_inc = mean(medincE),
         avg_age = mean(medageE)) %>%
  mutate(color = if_else(medageE < avg_age, GRAY9, ORANGE1))

```

Here is where we will do the mapping

```{r nailtedit1}

ggplot(pa_wide) + 
  geom_point(aes(x = medincE, y = medageE, color = color), size = 3) +
  scale_color_identity() +
  scale_x_continuous(limits = c(40000,100000), labels = scales::dollar_format()) +
  geom_hline(yintercept = pa_wide$avg_age, linetype = "longdash") + 
  geom_vline(xintercept = pa_wide$avg_inc, linetype = "longdash") +
  geom_point(x = pa_wide$avg_inc, y=pa_wide$avg_age, size=3.5) +
  labs(
    title = "Median Income vs. Median Age",
    subtitle = "PA Counties, 2020",
    x = "Median Income",
    y = "Median Age"
  )
  
  geom_point(aes(x=`Miles Driven`,y=`Cost Per Mile`, color = color), size = 3) +
  scale_color_identity() + 
  scale_y_continuous(breaks = seq(0,3,.5), limits = c(0,3),labels = scales::dollar_format()) + 
  scale_x_continuous(limits = c(0,4000), label = scales::comma) + 
  geom_hline(yintercept = avg_cost, linetype = "longdash") + 
  geom_point(x = avg_miles,y=avg_cost,size=3.5) + 
  geom_label(x = avg_miles, y = avg_cost, label = "AVG", hjust = 1.25, label.size = 0) +  
  labs(title = "Cost per mile by miles driven", x = "Miles driven per month", y = "Cost per mile") +
  coord_capped_cart(bottom = "right", left = "top")

```

Now let's explore the line graph

```{r data_fig209}
df <- read_csv(file.path("data", "FIG0209.csv")) %>% mutate(date = ymd(paste(Year, Month, 1)))
```

Here we will make the graph

```{r fig209}
pt <- ggplot(df, aes(x = date)) +
  geom_ribbon(aes(ymin = Min, ymax = Max), fill = GRAY9) +
  geom_line(aes(y = Avg), size = 2, color = GRAY4) +
  geom_point(aes(y = Avg), size = 1, data = df %>% slice(1), color = GRAY4) +
  geom_point(aes(y = Avg), size = 4, data = df %>% slice(n()), color = GRAY4) +
  geom_text(aes(y = Min), hjust = -0.1, vjust = -1.1, color = GRAY5, parse = T, label = sprintf("bold(MIN)"), data = df %>% slice(1)) +
  geom_text(aes(y = Avg), hjust = -0.1, vjust = -0.5, parse = T, label = "bold(AVG)", data = df %>% slice(1)) +
  geom_text(aes(y = Max), hjust = -0.1, vjust = 2, color = GRAY5, parse = T, label = "bold(MAX)", data = df %>% slice(1)) +
  geom_text(aes(y = Avg, label = paste0("bold(", Avg, ")")), parse = T, nudge_x = 15, data = df %>% slice(n())) +
  scale_y_continuous(breaks = seq(0, 40, 5), limit = c(0, 40)) +
  scale_x_date(
    date_labels = "%b", breaks = df$date, expand = c(0, 0),
    limits = c(ymd("2014-08-16", ymd("2015-10-01")))
  ) +
  labs(y = "Wait time (minutes)", x = NULL, title = "Passport control wait time", subtitle = "Past 13 months")

pt %>% 
  save_and_show_plot(width = 6, height = 4, "FIG0209.png")

```

Now the nailed it challenge is to re-create using longitudinal census data

```{r longitudinaldata}
years <- 2005:2019
names(years) <- years

philly_value <- map_dfr(years, ~{
  get_acs(
    geography = "county",
    variables = "B25077_001",
    state = "PA",
    county = "Philadelphia",
    year = .x,
    survey = "acs1"
  )
}, .id = "year")
```

Now for the nailed it graph! Rather than using the min and max fields from the previous example, use the estimate +/- MOE. Feel free to modify it so it looks BETTER than the sampler. 

```{r naileditlinegraph}


```


Let's try another example with some visual order. This is Figure 3.14 from SWD book


```{r Fig314}
library(tidyverse)
library(forcats)
library(lemon)
library(ggtext)
source("helper_functions.R")

# Read the CSV file from the specified path and load it into a dataframe
df <- read_csv(file.path("data", "FIG0314.csv")) %>%
  
  # Parse the 'value' column as a number and divide by 100 to convert percentages to proportions
  mutate(value = parse_number(value) / 100) %>%
  
  # Reorder the 'category' factor levels and reverse the order for a desired display order in plots
  mutate(category = fct_rev(fct_relevel(
    factor(category), # Convert 'category' column to a factor to control the order
    "Demonstration of results",    # Set the first level (highest priority) in the order
    "Content expertise",           # Followed by 'Content expertise'
    "Local knowledge",             # Continue setting levels in the desired order
    "National reputation",
    "Affordability of services",
    "Previous work together",
    "Colleague recommendation"     # The last level (lowest priority)
  ))) %>%
  
  # Create a new 'fill' column to assign colors based on the category
  mutate(fill = case_when(
    category == "Demonstration of results" ~ GRAY2,          # Assign GRAY2 color for 'Demonstration of results'
    category == "Affordability of services" ~ GRAY2,         # Assign GRAY2 color for 'Affordability of services'
    category == "Previous work together" ~ GRAY2,            # Assign GRAY2 color for 'Previous work together'
    TRUE ~ GRAY8                                             # Assign GRAY8 color for all other categories
  ))


pt <- ggplot(df, aes(x = category, y = value)) +
  geom_col(aes(fill = fill), width = .8) +
  scale_x_discrete() +
  scale_y_continuous(position = "right", limits = c(0, .8), labels = scales::percent) +
  scale_fill_identity() +
  coord_capped_flip(top = "both") +
  labs(
    caption = "Data source: xyz; indluces N number of survey respondents.\nNote that respondents were able to choose up to 3 options.",
    y = "% selecting given attribute",
    title = "**Demonstrating effectiveness** is most important consideration <br/>when selecting a provider <br />",
    subtitle = "In general, **what attributes are the most important** <br/>to you in selecting a service provider ? <br/> (Choose up to 3)"
  )+
  theme(
    plot.title = element_markdown(size = 14),   # Enables Markdown parsing for the title
    plot.subtitle = element_markdown(size = 10), # Enables Markdown parsing for the subtitle
    plot.caption = element_text(size = 8)
  )


pt %>% 
  save_and_show_plot(width = 6, height = 4, file_name = "FIG0314.png")







```


Apply those same principles to spruce up a boring old census bar chart! Best in show wins the award.
There is no award. We all win.

```{r bestinshow}
library(tidycensus)
library(tidyverse)

metros <-  get_acs(
  geography = "cbsa",
  variables = "DP03_0021P",
  summary_var = "B01003_001",
  survey = "acs1",
  year = 2019
) %>%
  slice_max(summary_est, n = 20)

metros %>%
  mutate(NAME = str_remove(NAME, "-.*$")) %>%
  mutate(NAME = str_remove(NAME, ",.*$")) %>%
  ggplot(aes(y = reorder(NAME, estimate), x = estimate)) + 
  geom_col()


metros %>%
  mutate(NAME = str_remove(NAME, "-.*$")) %>%
  mutate(NAME = str_remove(NAME, ",.*$")) %>%
  ggplot(aes(y = reorder(NAME, estimate), x = estimate)) + 
  geom_col() +  
  theme_minimal() + 
  labs(title = "Public transit commute share", 
       subtitle = "2019 1-year ACS estimates", 
       y = "", 
       x = "ACS estimate", 
       caption = "Source: ACS Data Profile variable DP03_0021P via the tidycensus R package") 



```


Spice it up here:

```{r}

```


Bees?? Can you update this beeswarm map following the tidytuesday example from the ppt?

```{r}
library(ggbeeswarm)

ny_race_income <- get_acs(
  geography = "tract", 
  state = "NY",  
  county = c("New York", "Bronx", "Queens", "Richmond", "Kings"),
  variables = c(White = "B03002_003", 
                Black = "B03002_004", 
                Asian = "B03002_006",
                Hispanic = "B03002_012"), 
  summary_var = "B19013_001",
  year = 2020
) %>%
  group_by(GEOID) %>%
  filter(estimate == max(estimate, na.rm = TRUE)) %>%
  ungroup() %>%
  filter(estimate != 0)

ggplot(ny_race_income, aes(x = variable, y = summary_est, color = summary_est)) +
  geom_quasirandom(alpha = 0.5) + 
  coord_flip() + 
  theme_minimal(base_size = 13) + 
  scale_color_viridis_c(guide = "none") + 
  scale_y_continuous(labels = label_dollar()) + 
  labs(x = "Largest group in Census tract", 
       y = "Median household income", 
       title = "Household income distribution by largest racial/ethnic group", 
       subtitle = "Census tracts, New York City", 
       caption = "Data source: 2016-2020 ACS")
```




